---
title: "Apurva Mahajan - Detroit"
output: html_document
date: "2025-04-22"
---

A demo notebook for Dow Jones News Fund data interns

During the week’s training, you will work on a project called Investigate Your Community to produce detailed story pitches relevant to their summer internship positions. The goal is to have the students research a topic relevant to your newsroom so they can get to work immediately at the internship. 

Here are some examples from last year’s DJNF data journalism training:

Hannah Rosenberger, at ASI Media/Counselor magazine, produced this analysis on name, image and likeness trends: https://hrosenberger.github.io/ASI-NIL-Project/

Shifra Dayak, an intern at the Connecticut Mirror, produced this analysis of environment spill data: https://shifradayak.github.io/DJNF_Merrill/ct_env_spills.html.

Michael Zhang, an intern at the Houston Landing, explored local election data: https://michaelzhangreporter.github.io/Houston_Campaign_Finance/


Here's what your newsroom suggested for a project:

Wondering if you could explore trends of our vaccination rates, especially considering we’re dealing with measles here.https://www.detroitnews.com/story/news/local/michigan/2025/04/24/michigan-headed-toward-worst-year-for-measles-in-decades-chief-medical-exec-says/83214966007/
 
This website explains the different Series numbers for the sets of vaccines: https://www.michigan.gov/mdhhs/adult-child-serv/childrenfamilies/immunizations/data-statistics/countyimmsreportcards 
 
If you or our intern have any questions, you can contact Lynn Sutfin at the Michigan Department of Health and Human Services at SutfinL1@michigan.gov. 




This notebook helps you get started

# Load software libraries
```{r}
#install.packages("tidyverse")
#install.packages("janitor")
#install.packages("readxl")
#install.packages("rvest")
library(tidyverse)
library(janitor)
library(readxl)
library(rvest)

```


# Load your data


```{r}

# 
# YOUR_DATA <- read.csv("../data/XXXXX.csv") |> 
#     clean_names()

data <- read_excel("../data/Report_card_All_data_20250313.xlsx") |> 
    clean_names()


```

Data details:


# Explore the data types
```{r}

#glimpse(YOUR_DATA)

glimpse(data)

```


# Data exploration: dates

```{r}
data |> 
   count(year) |> 
  arrange(desc(n))
```




# Questions 

Write down questions you want answered:



# Data Details
Source:
https://www.michigan.gov/mdhhs/adult-child-serv/childrenfamilies/immunizations/data-statistics/countyimmsreportcards

Documentation:
https://www.michigan.gov/mdhhs/-/media/Project/Websites/mdhhs/Adult-and-Childrens-Services/Children-and-Families/Immunization-Information/LHD/Immunization-Report-Cards/Report_Card_Defs__FAQs_2021.pdf?rev=93e4ac10b66c4c10abcd3c02a6b55982&hash=EA6122E84CB516E9E6D335CC73BDB5CF

