---
title: "Apurva Mahajan - Detroit"
output: html_document
date: "2025-04-22"
---

A demo notebook for Dow Jones News Fund data interns

During the week's training, you will work on a project called Investigate Your Community to produce detailed story pitches relevant to their summer internship positions. The goal is to have the students research a topic relevant to your newsroom so they can get to work immediately at the internship.

Here are some examples from last year's DJNF data journalism training:

Hannah Rosenberger, at ASI Media/Counselor magazine, produced this analysis on name, image and likeness trends: <https://hrosenberger.github.io/ASI-NIL-Project/>

Shifra Dayak, an intern at the Connecticut Mirror, produced this analysis of environment spill data: <https://shifradayak.github.io/DJNF_Merrill/ct_env_spills.html>.

Michael Zhang, an intern at the Houston Landing, explored local election data: <https://michaelzhangreporter.github.io/Houston_Campaign_Finance/>

Here's what your newsroom suggested for a project:

Wondering if you could explore trends of our vaccination rates, especially considering we're dealing with measles here.https://www.detroitnews.com/story/news/local/michigan/2025/04/24/michigan-headed-toward-worst-year-for-measles-in-decades-chief-medical-exec-says/83214966007/

This website explains the different Series numbers for the sets of vaccines: <https://www.michigan.gov/mdhhs/adult-child-serv/childrenfamilies/immunizations/data-statistics/countyimmsreportcards>

If you or our intern have any questions, you can contact Lynn Sutfin at the Michigan Department of Health and Human Services at [SutfinL1\@michigan.gov](mailto:SutfinL1@michigan.gov){.email}.

Background info: There have been 9 reported measles cases this year in Michigan, in Ingham, Kent, Macomb, Marquette, Montcalm and Oakland Counties.

This notebook helps you get started

# Load software libraries

```{r}
#install.packages("tidyverse")
#install.packages("janitor")
#install.packages("readxl")
#install.packages("rvest")
library(tidyverse)
library(janitor)
library(readxl)
library(rvest)
library(zoo)

```

# Load your data

```{r}

# The report card data provides estimates for immunization coverage at the county level for vaccines recommended for children (19-35 months) and adolescent (13-17-year-olds)

# load in vaccine report card data
report_card_data <- read.csv("./data/Report_card_All_data_20250313.csv") |> 
    clean_names()

# Since the time for this data is by quarter, I'm going to create a column based on an index so I can plot this and analyze by time. 

report_card_data <- report_card_data %>% 
  mutate(
     quarter_num = as.numeric(sub("Q", "", quarter)),
     quarter_index = year + (quarter_num - 1) / 4,
    quarter_date = make_date(year, (quarter_num - 1) * 3 + 1, 1)
  )

# Now I'm going to create a column that calculates the percent change from 2010 Q4 to 2025 Q4. What counties have seen the highest changes in Hepatitis A vaccination coverage.

report_card_data %>% 
  group_by(cntyname) %>% 
  filter(quarter_index %in% c(2014.75, 2024.75)) %>% 
  select(cntyname, quarter_index, covg_hep_a) %>% 
  pivot_wider(
    names_from = quarter_index,
    values_from = covg_hep_a,
    names_prefix = "Q_"
  ) %>% 
  mutate(pct_change_2014_2024 = 100 * (Q_2024.75 - Q_2014.75) / Q_2014.75) %>% arrange(pct_change_2014_2024)

# now lets look at the 4313314 vaccine, which is: 
  # "4 or more doses of DTaP/DTP/DT, 3 or more doses of Polio, 1 or more dose of MMR, 3 or more doses of Hib, 3 ore more doses of HepB, 1 or more dose of Varicella, 4 or more doses of PCV" 

report_card_data %>% 
  group_by(cntyname) %>% 
  filter(quarter_index %in% c(2014.75, 2024.75)) %>% 
  select(cntyname, quarter_index, covg4313314) %>% 
  pivot_wider(
    names_from = quarter_index,
    values_from = covg4313314,
    names_prefix = "Q_"
  ) %>% 
  mutate(pct_change_2014_2024 = 100 * (Q_2024.75 - Q_2014.75) / Q_2014.75) %>% 
  arrange(pct_change_2014_2024)

# Coverage for the 4313314 vaccine decreased for all but three counties in Michigan. 
  


# Checking out the basics of the database. 
# coverage refers to the percent of the population who have the immunization. 


# MMR coverage 
report_card_data %>% 
  filter(year == 2024, quarter == "Q4") %>% 
  group_by(cntyname, year) %>%
  summarize(covgmmr) %>% 
  arrange(covgmmr)

# Hepatitis A vaccination coverage for 2024
report_card_data %>% 
  filter(year == 2024, quarter == "Q4") %>% 
  group_by(cntyname, year) %>%
  summarize(covg_hep_a) %>% 
  arrange(covg_hep_a)

# dTaP coverage
report_card_data %>% 
  filter(year == 2024, quarter == "Q4") %>% 
  group_by(cntyname, year) %>%
  summarize(covg_d_tap) %>% 
  arrange(covg_d_tap)

# pcv7 coverage

report_card_data %>% 
  filter(year == 2024, quarter == "Q4") %>% 
  group_by(cntyname, year) %>%
  summarize(covg_pcv7) %>% 
  arrange(covg_pcv7)

# 

ls(data)


# 
report_card_data %>% 
  filter(cntyname == "Wayne") %>% 
  ggplot() +
  geom_line (aes(x=quarter_index, y=covg_d_tap))
```

```{r}

# Let's look at MMR vaccination rates as of Dec. 31, 2024 (end of 2024 Q4)

report_card_data %>% 
  filter(
    year == "2024",
    quarter == "Q4"
  ) %>% 
  select(year, quarter, cntyname, covgmmr) %>% 
  arrange(covgmmr) %>% 
  slice(1:10)

# Oscoda County had the lowest MMR vaccination rates, followed by Keweenaw, Gladwin, Mackinac and Salinac. None of these counties saw measles cases in 2025.





```

Data details:

# Explore the data types

```{r}

#glimpse(YOUR_DATA)

glimpse(data)

```

# Data exploration: dates

```{r}
data |> 
   count(year) |> 
  arrange(desc(n))
```

# Questions

Write down questions you want answered:

# Potential sources:

# Thrishika Bala, epidemiologist for the state of Michigan

# Data Details

Source: <https://www.michigan.gov/mdhhs/adult-child-serv/childrenfamilies/immunizations/data-statistics/countyimmsreportcards>

Documentation: <https://www.michigan.gov/mdhhs/-/media/Project/Websites/mdhhs/Adult-and-Childrens-Services/Children-and-Families/Immunization-Information/LHD/Immunization-Report-Cards/Report_Card_Defs__FAQs_2021.pdf?rev=93e4ac10b66c4c10abcd3c02a6b55982&hash=EA6122E84CB516E9E6D335CC73BDB5CF>
